# Installing R Markdown and latex {#install-rm}

Installing R markdown on your computer is pretty straight forward and should be painless. If you're getting started with R markdown we suggest that you use RStudio but of course RStudio is not required and there are other options available. You will also need to install some additional software if you want to render your R markdown documents to PDF format. 

This guide assumes you have already installed [R][r-proj] and an IDE ([RStudio IDE][rstudio] or [VSCode][vscode]. An IDE is not required but recommended, because it makes it easier to work with R Markdown. If you don't have RStudio IDE installed, you will also have to install [Pandoc][pandoc]. If you have RStudio installed there is no need to install Pandoc separately because it's bundled with RStudio. Next you can install the `rmarkdown` package in RStudio using the following code:

```{r, eval = FALSE}
# Install from CRAN
install.packages('rmarkdown', dep = TRUE)
```

The `dep = TRUE` argument will also install a bunch of additional R packages on which rmarkdown depends.
  
If you want to generate PDF output, you will need to install LaTeX. For R Markdown users who have not installed LaTeX before, we recommend that you install [TinyTeX][tinytex]. You can install TinyTex from within RStudio using the following code:

```{r, eval = FALSE}
install.packages('tinytex')
tinytex::install_tinytex()  # install TinyTeX
```

TinyTeX is a lightweight, portable, cross-platform, and easy-to-maintain LaTeX distribution. The R companion package tinytex can help you automatically install missing LaTeX packages when compiling LaTeX or R Markdown documents to PDF.

## MS Windows

An alternative option would be to install MiKTeX instead. You can download the latest distribution of [MiKTeX][miktex]. Installing MiKTeX is pretty straight forward, but it can sometimes be a pain to get it to play nicely with RStudio. If at all possible we recommend that you use TinyTex. 

## Mac OSX

If for some reason TinyTeX does not work on your Mac computer then you can try to install MacTeX instead. You can download the latest version of [MacTeX here][mactex].

## Linux

An alternative to TinyTex on linux would be to use a full fledge distribution of latex such as [TexLive][texlive]


```{r links, child="_links.md"}
```


